---
title: "Experiment 1 analysis and visualizations"
author: "MH Tessler"
date: "1/31/2020"
output: github_document
---

```{r libraries}
library(tidyverse)
library(jsonlite)
library(viridis)
library(tidyboot)
# library(brms)
# library(lme4)
# library(lmerTest)
library(ggridges)
library(ggstance)
#library(rwebppl)
library(knitr)
library(ggthemes)
theme_set(theme_few())
```


```{r read data from csv}
df.trials <- read_csv("../data/expt1/genex-cogsci_expt1-trials.csv")
df.subject <- read_csv("../data/expt1/genex-cogsci_expt1-participant_information.csv")
df.attention <- read_csv("../data/expt1/genex-cogsci_expt1-sound_check.csv")
```

# Attention checks and other exclusions

## Multiple Submitters

Some participants were able to complete the experiment multiple times. 
Here, I grab only the data from their first submission.

```{r multiple submitters}
# calculate # of submissions per work
df.trials %>% 
  group_by(workerid, condition) %>% 
  summarize(n = n() / 6) %>% # expt consists of 6 trials
  ungroup() %>% 
  group_by(workerid) %>% 
  summarize(n = sum(n)) -> submissions_per_worker

# how many submissions were by people who did it more than once?
submissions_per_worker %>% 
  filter(n > 1) %>% 
  ungroup() %>% 
  summarize(sum(n))

# how many submissions in total?
submissions_per_worker %>% 
  ungroup() %>% 
  summarize(total_subj = sum(n))

# grab first instance of participant completing a n expt
## grab: condition, kind_type, kind_label, feature_label, and response
df.trials.removeDuplicateParticipants <- df.trials %>% 
  group_by(workerid, exptid, trial_type, trial_num) %>%
  summarize(condition = first(condition),
    kind_type = first(kind_type), 
    kind_label = first(kind_label),
    feature_label = first(feature_label),
    response = first(response))
```

Test to see "multiple submitters" exclusions are reflected in the number of memory check trials that completed correctly and the number of trials they completed.

```{r test multiple submitters}
# check to see exclusions are reflected in the number of memory check trials that completed correctly
## should be less than or equal to 3
df.trials.removeDuplicateParticipants %>% 
  filter(trial_type == 'id') %>%
  group_by(workerid, condition) %>%
  summarize(n_correct = sum(response)) %>%
  ungroup() %>%
  group_by(n_correct) %>%
  count()

# number of memory check trials completed by each participant
## should always be 3
df.trials.removeDuplicateParticipants %>% 
  filter(trial_type == 'id') %>%
  group_by(workerid) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(n) %>%
  count()
```

## Memory check

Proportion of correct IDs broken down by item. 
Artifact is a little harder to discriminate in memory check.

```{r}
df.trials.removeDuplicateParticipants %>%
  filter(trial_type == 'id') %>%
  group_by(kind_type) %>%
  summarize(prop_correct = sum(response) / n()) %>%
  kable()
```

Remove participants who do not answer all 3 correctly.

```{r memory check}
df.id.catch <- df.trials.removeDuplicateParticipants %>%
  filter(trial_type == 'id') %>%
  group_by(workerid, condition) %>%
  summarize(n_correct = sum(response), # no of memory check trials answered correctly
            pass = n_correct == 3) # must answer all 3 to be included

df.trials.filtered.mem <- df.trials.removeDuplicateParticipants %>%
  left_join(., df.id.catch) %>% 
  filter(pass)
```


```{r test memory check}
df.trials.filtered.mem %>%
  filter(trial_type == 'id') %>%
  group_by(workerid) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(n) %>%
  count()
```

## Sound check

"skyscraper" was the distractor word on the screen that a participant would be drawn to if they are not actually reading the text. we filter based on responses of "skyscraper" or a misspelling.

```{r sound check}
# Workers pass sound check
df.attention %>%
  group_by(workerid, exptid) %>%
  summarize(test_word = first(test_word),
  response = first(response)) %>%
  mutate(response = tolower(response)) %>%
  rowwise() %>%
  mutate(pass_sound = !(response %in% c("skyscraper", "skycraper"))) -> df.speaker.fail

# how many fail?
sum(!df.speaker.fail$pass_sound)

## check what participants wrote when they got it wrong
# df.speaker.fail %>% filter(!pass_sound) %>% View()
df.trials.filtered <- df.trials.filtered.mem %>%
  left_join(., df.speaker.fail %>% select(workerid,pass_sound)) %>% 
  filter(pass, pass_sound)
```



## Number of Participants per condition

```{r}
length(unique(df.trials.filtered$workerid)) # data frame after exlcusions

length(unique(df.trials$workerid)) # original data frame

df.trials.filtered %>% 
  select(workerid, condition, kind_type, trial_type) %>%
  filter(kind_type == "bird", trial_type == "response") %>%
  group_by(condition) %>%
  count() %>%
  kable()
```


# Analysis and Visualization

```{r relevel}
# grab main trail responses and relevel condition factor
df.trials.filtered.recoded <- df.trials.filtered %>%
  filter(trial_type == 'response') %>%
  mutate(condition = factor(condition,
                            levels = c("accidental",
                                       "2accidental",
                                       "3accidental",
                                       "4accidental",
                                       "pedagogical",
                                       "2pedagogical",
                                       "3pedagogical",
                                       "4pedagogical",
                                       "generic",
                                       "pedageneric"),
                            labels = c("1x Accidental",
                                       "2x Accidental",
                                       "3x Accidental",
                                       "4x Accidental",
                                       "1x Pedagogical",
                                       "2x Pedagogical",
                                       "3x Pedagogical",
                                       "4x Pedagogical",
                                       "Generic",
                                       "Generic \n+ 1x Pedagogical")))

## check: Number of responses per subject
df.trials.filtered.recoded %>%
  group_by(workerid) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  group_by(n) %>%
  count()
```

Compute bootstrapped 95% confidence intervals, by condition and by-itemXcondition
```{r CIs}
## 95 CIs
df.bs <- df.trials.filtered.recoded %>%
  group_by(condition) %>%
  tidyboot_mean(column = response) %>%
  ungroup() %>%
  mutate(condition = factor(condition))

df.bs.item <- df.trials.filtered.recoded %>%
  group_by(condition, kind_type, feature_label) %>%
  tidyboot_mean(column = response) %>%
  ungroup() %>%
  mutate(condition = factor(condition),
         kind_type = factor(kind_type, levels = c("bird", "flower", "artifact")))
```

## Ridge distribution plot (Figure 3A)

```{r}
ggplot(df.trials.filtered.recoded, 
       aes(x = response, y = condition, fill = ..x..)) +
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.9,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.3,
    rel_min_height = 0.01, gradient_lwd = 1,
    stat = 'binline', bins = 25, draw_baseline = F
  ) +
  geom_linerangeh(data = df.bs,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper,
        y = as.numeric(condition)+0.4),
    size = 1.25, color = 'black')+
  geom_vline(data = df.bs %>% filter(condition == "Generic"), 
             linetype = 2, alpha = 0.4,
             aes(xintercept = ci_lower))+
  geom_vline(data = df.bs %>% filter(condition == "Generic"), 
             linetype = 2, alpha = 0.4,
             aes(xintercept = ci_upper))+
  geom_point(data = df.bs,
    inherit.aes = F,
    aes(x = mean,
        y = as.numeric(condition)+0.4),
    size = 3, color = 'black', shape = 3)+
  scale_x_continuous(expand = c(0.01, 0), 
                     limits = c(0, 1.05), 
                     breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_viridis(name = "Implied Prevalence", option = "D",
                     breaks = c(0, 1)) +
  guides(fill = F)+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(hjust = 0.5, vjust = 0))+
  labs(x = "Probability of Future Instance having Property")

# ggsave("../paper/figs/genex-expt1.pdf", width = 5, height = 5)
```

## By-item bar plot (Figure 4A)

```{r}
bar.width = 0.7

df.bs.item %>%
  ggplot(., aes(x = condition, fill = kind_type, 
                y = mean, ymin = ci_lower,
                ymax = ci_upper)) +
  geom_hline(data = df.bs %>% filter(condition == "Generic"),
             linetype = 2, alpha = 0.4,
             aes(yintercept = ci_lower))+
  geom_hline(data = df.bs %>% filter(condition == "Generic"),
             linetype = 2, alpha = 0.4,
             aes(yintercept = ci_upper))+
  geom_col(color = 'black', width = bar.width, position = position_dodge(bar.width),
           alpha = 0.5)+
  geom_linerange( position = position_dodge(bar.width))+
  coord_flip()+
   guides(fill = guide_legend(reverse = T))+
  labs(y = "Probability of Future Instance having Property", x = "",
       fill = "item")+
  theme(legend.position = 'bottom')

# ggsave("../paper/figs/genex10conditions_byItem_bars_reordered.pdf",
#        width = 5, height = 4.5)

```


# Other plots

## Bars

```{r}
df.bs %>%
  ggplot(., aes(x = condition, y = mean, ymin = ci_lower,
                ymax = ci_upper)) +
  geom_col(color = 'black', fill = 'white')+
  geom_linerange()+
  coord_flip()
```

## Histograms

```{r}
df.trials.filtered.recoded %>%
  ggplot(., aes(x = response))+
  geom_histogram()+
  facet_wrap(~condition, nrow = 2)+
  scale_x_continuous(breaks = c(0, 1))
```


## By-item plots

### by-item Histograms

```{r}
df.trials.filtered.recoded %>%
  ggplot(., aes(x = response))+
  geom_histogram()+
  facet_grid(feature_label~condition)
```

### by-item ridge plots

```{r}
df.trials.filtered.recoded %>%
  mutate(
    kind_type = factor(kind_type, levels = c("bird", "flower", "artifact"))
    ) %>%
  ggplot(., 
       aes(x = response, y = condition, fill = ..x..)) +
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.9,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.3,
    rel_min_height = 0.01, gradient_lwd = 1,
    stat = 'binline', bins = 25, draw_baseline = F
  ) +
  geom_linerangeh(data = df.bs.item,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper,
        y = as.numeric(condition)+0.4),
    size = 1.25, color = 'black')+
  geom_vline(data = df.bs.item %>% filter(condition == "Generic"), 
             #inherit.aes = F, 
             linetype = 2, alpha = 0.4,
             aes(xintercept = ci_lower))+
  geom_vline(data = df.bs.item %>% filter(condition == "Generic"), 
             #inherit.aes = F, 
             linetype = 2, alpha = 0.4,
             aes(xintercept = ci_upper))+
  geom_point(data = df.bs.item,
    inherit.aes = F,
    aes(x = mean,
        y = as.numeric(condition)+0.4),
    size = 3, color = 'black', shape = 3)+
  scale_x_continuous(expand = c(0.01, 0), 
                     limits = c(0, 1.02), 
                     breaks = c(0, 1)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_viridis(name = "Implied Prevalence", option = "D",
                     breaks = c(0, 1)) +
  guides(fill = F)+
  facet_wrap(~kind_type, nrow = 1)+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(hjust = 0.5, vjust = 0))+
  labs(x = "Probability of Future Instance having Property")

ggsave("../paper/figs/genex10conditions_byItem_reordered.pdf",
       width = 10, height = 4.5)

```


## By-trial num plots

### By-trial num histograms

```{r}
df.trials.filtered.recoded %>%
  ggplot(., aes(x = response))+
  geom_histogram()+
  facet_grid(trial_num~condition)
```


### By-trial num bar plots

```{r}
df.bs.trial_num <- df.trials.filtered.recoded %>%
  group_by(condition, trial_num) %>%
  tidyboot_mean(column = response)
```

```{r}
bar.width = 0.7

df.bs.trial_num %>%
  mutate(trial_num = factor(trial_num)) %>%
  ggplot(., aes(x = condition, fill = trial_num, 
                y = mean, ymin = ci_lower,
                ymax = ci_upper)) +
  geom_hline(data = df.bs %>% filter(condition == "Generic"),
             linetype = 2, alpha = 0.4,
             aes(yintercept = ci_lower))+
  geom_hline(data = df.bs %>% filter(condition == "Generic"),
             linetype = 2, alpha = 0.4,
             aes(yintercept = ci_upper))+
  geom_col(color = 'black', width = bar.width, position = position_dodge(bar.width),
           alpha = 0.5)+
  geom_linerange( position = position_dodge(bar.width))+
  coord_flip()+
   guides(fill = guide_legend(reverse = T))+
  labs(y = "Probability of Future Instance having Property", x = "",
       fill = "item")+
  theme(legend.position = 'bottom')
```


## Subset of conditions

### Ridgeplots: Pedagogical and Generic only

```{r}
generic.and.pedagogical <-  c("1x Pedagogical",
                              "2x Pedagogical",
                              "3x Pedagogical",
                              "4x Pedagogical",
                              "Generic",
                              "Generic \n+ 1x Pedagogical")



df.trials.filtered.recoded.ped <- df.trials.filtered.recoded %>% 
  filter(condition %in% generic.and.pedagogical) %>%
  ungroup() %>% 
  mutate(condition = factor(condition, levels = generic.and.pedagogical))

df.bs.ped <- df.bs %>% 
  filter(condition %in% generic.and.pedagogical) %>%
  ungroup() %>% 
  mutate(condition = factor(condition, levels = generic.and.pedagogical))

ggplot(df.trials.filtered.recoded.ped, 
       aes(x = response, y = condition, fill = ..x..)) +
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.95,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.3,
    rel_min_height = 0.01, gradient_lwd = 1,
    stat = 'binline', bins = 25, draw_baseline = F
  ) +
  geom_linerangeh(data =df.bs.ped,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper, 
        y = as.numeric(condition)+0.4),
    size = 1.25, color = 'black')+
  geom_point(data = df.bs.ped,
    inherit.aes = F,
    aes(x = mean,
        y = as.numeric(condition)+0.4),
    size = 4, color = 'black', shape = 3)+
  scale_x_continuous(expand = c(0.01, 0), 
                     limits = c(0, 1.02), 
                     breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_viridis(name = "Implied Prevalence", option = "D",
                     breaks = c(0, 1)) +
  guides(fill = F)+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(hjust = 0.5, vjust = 0))+
  labs(x = "P(Feature | Category)")

# ggsave("~/projects/generic-interpretation/posters/genex-pilots_ped.png",
#        width = 5, height =4 )
```

### Ridgeplots: Accidental and Generic only

```{r}
generic.and.accidental <-  c("1x Accidental",
                              "2x Accidental",
                              "3x Accidental",
                              "4x Accidental",
                              "Generic",
                              "Generic \n+ 1x Pedagogical")

df.trials.filtered.recoded.acc <- df.trials.filtered.recoded %>% 
  filter(condition %in% generic.and.accidental) %>%
  ungroup() %>% 
  mutate(condition = factor(condition, levels = generic.and.accidental))

df.bs.acc <- df.bs %>% 
  filter(condition %in% generic.and.accidental) %>%
  ungroup() %>% 
  mutate(condition = factor(condition, levels = generic.and.accidental))

ggplot(df.trials.filtered.recoded.acc, 
       aes(x = response, y = condition, fill = ..x..)) +
  geom_density_ridges_gradient(
    jittered_points = T, alpha = 0.8, scale = 0.95,
    position = position_points_jitter(width = 0.01, height = 0),
    point_shape = '|', point_size = 2.5, point_alpha = 0.3,
    rel_min_height = 0.01, gradient_lwd = 1,
    stat = 'binline', bins = 25, draw_baseline = F
  ) +
  geom_linerangeh(data =df.bs.acc,
    inherit.aes = F,
    aes(xmin = ci_lower, xmax = ci_upper, 
        y = as.numeric(condition)+0.4),
    size = 1.25, color = 'black')+
  geom_point(data = df.bs.acc,
    inherit.aes = F,
    aes(x = mean,
        y = as.numeric(condition)+0.4),
    size = 4, color = 'black', shape = 3)+
  scale_x_continuous(expand = c(0.01, 0), 
                     limits = c(0, 1.02), 
                     breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_viridis(name = "Implied Prevalence", option = "D",
                     breaks = c(0, 1)) +
  guides(fill = F)+
  theme(axis.title.y = element_blank(),
        axis.title.x = element_text(hjust = 0.5, vjust = 0))+
  labs(x = "P(Feature | Category)")

# ggsave("~/projects/generic-interpretation/posters/genex-pilots_acc.png",
#        width = 5, height = 4)
```

# Regression

### Forward difference


```{r}

my.forward.diff = matrix(
  c(7/8, -1/8, -1/8, -1/8, -1/8, -1/8, -1/8, -1/8, 
    6/8, 2/8, -(4/3)/8, -(4/3)/8, -(4/3)/8, -(4/3)/8, -(4/3)/8, -(4/3)/8, 
    5/8, 3/8, 3/8, -11/40, -11/40, -11/40, -11/40, -11/40,
    1/2, 1/2, 1/2, 1/2, -1/2, -1/2, -1/2, -1/2,
    11/40, 11/40, 11/40, 11/40, 11/40, -3/8, -3/8, -5/8,
    (4/3)/8, (4/3)/8, (4/3)/8, (4/3)/8, (4/3)/8, (4/3)/8, -2/8, -6/8,
    1/8, 1/8, 1/8, 1/8, 1/8, 1/8, 1/8, -7/8),
      ncol = 7)


contrasts(df.trials.filtered.recoded$condition) <- my.forward.diff


rs.lm <- lmer(response ~ 
       condition + (1 | workerid) + (1 | singular), data = df.trials.filtered.recoded)

summary(rs.lm)
```

### Dummy coding

```{r}
df.trials.filtered.recoded.genDummy <- df.trials.filtered.recoded %>%
  mutate(condition = factor(condition,
                            levels = c("Generic",
                                       "1x Accidental",
                                       "1x Pedagogical",
                                       "2x Accidental",
                                       "2x Pedagogical",
                                       "3x Accidental",
                                       "3x Pedagogical",
                                       "4x Pedagogical",
                                       "4x Accidental",
                                       "Generic \n+ 1x Pedagogical")))

contrasts(df.trials.filtered.recoded.genDummy$condition) 

rs.lm.dummy <- lmer(response ~ condition + (1 | workerid) + 
                      (1 | kind_type), data = df.trials.filtered.recoded.genDummy)

summary(rs.lm.dummy)
```


```{r}
rs.brm.dummy <- brm(response ~ condition + (1 | workerid) + 
                      (1 | singular), 
                    data = df.trials.filtered.recoded.genDummy,
                    chains = 2, cores = 2)
summary(rs.brm.dummy)
```

```{r}
rs.brm.dummy <- brm(response ~ condition + (1 | workerid) + 
                      (1 | singular), 
                    data = df.trials.filtered.recoded.genDummy,
                    chains = 2, cores = 2)
summary(rs.brm.dummy)
```

```{r}
rs.brm.dummy.beta <- brm(response ~ condition + (1 | workerid) + 
                      (1 | singular), 
                    data = df.trials.filtered.recoded.genDummy %>%
                      mutate(response = ifelse(response == 1, 0.999, 
                                               ifelse(response == 0, 0.001, response))),
                    chains = 2, cores = 2, family = Beta()
                    )
summary(rs.brm.dummy.beta)
```


```{r}

df.trials.filtered.recoded.genDummy2 <- df.trials.filtered.recoded.genDummy %>%
  mutate(condition_flat = tolower(gsub(" ", "", as.character(condition), fixed = TRUE)),
         condition_flat = gsub("\n", "", condition_flat),
         condition_flat = gsub("\\+", "", condition_flat)
         ) %>%
  mutate(condition_flat = factor(condition_flat,
                            levels = c("generic",
                                       "1xaccidental",
                                       "1xpedagogical",
                                       "2xaccidental",
                                       "2xpedagogical",
                                       "3xaccidental",
                                       "3xpedagogical",
                                       "4xpedagogical",
                                       "4xaccidental",
                                       "generic1xpedagogical")))


rs.brm.dummy.betaInfl <- brm(response ~ condition_flat + (1 | workerid) + 
                      (1 | kind_type), data = df.trials.filtered.recoded.genDummy2,
                    chains = 3, cores = 3, iter = 2000,
                    save_all_pars = TRUE,
                    family = zero_one_inflated_beta(),
                    control = list(adapt_delta = 0.98)
                    )

summary(rs.brm.dummy.betaInfl)
```


# Output derivative files

Output "generic only" condition separately for compilation with Expt. 2 data. 
Also output filtered data for model.

```{r eval = F}
df.trials.filtered %>% 
  filter(condition == "generic") -> df.trials.filtered.generic

# write_csv(df.trials.filtered.generic, path = "../data/expt2/genex-cogsci_expt1-genericOnly.csv")
# write_csv(df.trials.filtered, path = "../data/expt1/genex-cogsci_expt1-filtered.csv")
```

